Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-06 Manuscript/manuscript.Rmd
Checked !!!, cleanup, questions remain

[!!! Separated paragraphs]

@@

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC029) [@rao20143d]. The chromosomal coordinates of topologically associating domains throughout the genome were obtained using the arrowhead algorithm from the Juicer toolbox provided by the Aiden Laboratory (https://github.com/aidenlab/juicer) [@durand2016juicebox]. Using the juicer toolbox, we were able to call TADs [??? You called TADs yourself, correct? !!!Yes, using the arrowhead algoirthm from the juicer toolbox] on [??? how many; !!! the first 29 hic matrices (HIC001 to HIC029); should I provide more detail?] contact matrices at 5 kb, 25 kb, 50 kb, and 100 kb resolutions. Identified TAD boundaries were represented by their location on the linear genome (GRCh37/hg19 human genome assembly), including start and end coordinates for chromosomes 1 through 22. The start and end coordinates of all hierarchical TADs were concatenated, sorted, and unique TAD boundary points were obtained.

@@

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For each training set, an $l_{1}$ norm penalized regularization (LASSO) was performed. The $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10.

[??? To be modified for using LASSO only; !!!removed variable reduction paragraph and combined it above]

[??? Where is the resampling techniques come?; !!!mentioned it below]
### RE-SAMPLING TECHNIQUES

@@

4. **SMOTE** (Synthetic Minority Over-Sampling Technique) [@chawla2002smote]: This method incorporated both random under- and over-sampling. Under-sampling is performed without replacement from the majority class, while over-sampling is performed by creating new synthetic observations using the minority class. The user is able to specify the parameters denoting the percent of over- and under-sampling from the two classes. We used SMOTE to create perfectly balanced data. The `DMwR` R package (version 3.4.2) [??? Add version number; !!!added version] was used to perform SMOTE.

@@

The coordinates of the corresponding 5kb resolution contact maps resulted in a total of 44948 genomic bins [??? across all chromosomes? If I do 44948*5000, I get 224740000, which is 1 billion bases less than the size of the genome. What is going on?; !!!This is only CHR1 for right now] (Table 2). There were only 3083 (6.9%) bins that overlapped with a flanked TAD boundary [??? What is flanked TAD boundary? Why do we care?]. An increase in resolution resulted in an overall decrease in the number of genomic bins, as well as a slighly less degree of class imbalance. The percentage of imbalance for each of the other resolutions were found to be  14.0%, 13.7%, and 12.2% for 25 kb, 50kb, and 100kb resolutions respectively. [??? Add exact numbers of class imbalance for each resolution; !!! added exact numbers] (Supplementaty Table 2). [??? Do not understand. I understand the ratio of #minority/#majority, but what % tells?; !!!was overally average of class imbalance for the 4 data resolutions; removed sentence; doesn't seem relevant].  After applying each resampling technique to the training sets for each data resolution, we obtained perfectly balanced classes (Figure 4). That is, the ratio of bins that overlapped with flanked TAD boundaries to bins that did not was 1:1.

@@

We first assessed performances on random forest models built on fully unbalanced training sets across each resolution, for each predictor type (Figure 5) [??? Why not distance?; !!!included a  more desciptive plot with distance]. When looking at the the conventional metrics used for assessment (accuracy and AUC), models that were not corrected for class imbalance appeared to perform reasonably well.  However, when we considered the F1-score and MCC, we observed a significant degredation in performance across both resolution and predictor type [??? Describe, why is it so?; !!!described]. This was due in part to each models' inability to correctly classify the genomic bins that overlapped with flanked TAD boundaries, as recognized by the small number of true positive cases (Supplementary Table 3). Even without correcting for the class imbalance, we found that, within predictor types, models with distance type predictors consistently outperformed the other two predictor types across all performance metrics.

>>>>

@@

Replicate Hi-C data for the GM12878 cell line were downloaded from GEO GSE63525 (Experiments HIC001 to HIC029) [@rao20143d]. The chromosomal coordinates of topologically associating domains throughout the genome were obtained using the Arrowhead algorithm from the Juicer toolbox provided by the Aiden Laboratory (https://github.com/aidenlab/juicer) [@durand2016juicebox]. [??? how many; !!! the first 29 hic matrices (HIC001 to HIC029); should I provide more detail? ??? TBD, why 29?] contact matrices at 5 kb, 25 kb, 50 kb, and 100 kb resolutions. Identified TAD boundaries were represented by their location on the linear genome (GRCh37/hg19 human genome assembly), including start and end coordinates for chromosomes 1 through 22. The start and end coordinates of all hierarchical TADs were concatenated, sorted, and unique TAD boundary points were obtained.

@@

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the majority and minority sets. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For each training set, an $l_{1}$ norm penalized regularization (LASSO) was performed [??? what for?]. The $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10.

### RE-SAMPLING TECHNIQUES

@@

4. **SMOTE** (Synthetic Minority Over-Sampling Technique) [@chawla2002smote]: This method incorporated both random under- and over-sampling. Under-sampling is performed without replacement from the majority class, while over-sampling is performed by creating new synthetic observations using the minority class. The user is able to specify the parameters denoting the percent of over- and under-sampling from the two classes. We used SMOTE to create perfectly balanced data. The `DMwR` R package (version 3.4.2) was used to perform SMOTE.

@@

The coordinates of the corresponding 5kb resolution contact maps resulted in a total of 44948 genomic bins [??? To be modified for the whole genome] (Table 2). There were only 3083 (6.9%) bins that overlapped with a flanked TAD boundary [??? What is flanked TAD boundary? Why do we care?]. An increase in resolution resulted in an overall decrease in the number of genomic bins, as well as a slighly less degree of class imbalance. The percentage of imbalance for each of the other resolutions were found to be  14.0%, 13.7%, and 12.2% for 25 kb, 50kb, and 100kb resolutions respectively. [??? Add exact numbers of class imbalance for each resolution; !!! added exact numbers] (Supplementaty Table 2). [??? Do not understand. I understand the ratio of #minority/#majority, but what % tells?; !!!was overally average of class imbalance for the 4 data resolutions; removed sentence; doesn't seem relevant ??? We should rewrite using the ratio of #minority/#majority].  After applying each resampling technique to the training sets for each data resolution, we obtained perfectly balanced classes (Figure 4). That is, the ratio of bins that overlapped with flanked TAD boundaries to bins that did not was 1:1.

@@

We first assessed performances on random forest models built on fully unbalanced training sets across each resolution, for each predictor type (Figure 5) [??? Why not distance?; !!!included a  more desciptive plot with distance ??? Where? Don't see it on Figure 5, which exists in two versions (?)]. When looking at the the conventional metrics used for assessment (accuracy and AUC), models that were not corrected for class imbalance appeared to perform reasonably well.  However, when we considered the F1-score and MCC, we observed a significant degradation in performance across both resolution and predictor type. This was due in part to each models' inability to correctly classify the genomic bins that overlapped with flanked TAD boundaries (true positives, Supplementary Table 3). Even without correcting for the class imbalance, we found that, within predictor types, models with distance type predictors consistently outperformed the other two predictor types across all performance metrics.
